### 2024-11-13 code/R/TimeVaryingDetection_mrmr.Rmd
initial commit

---
title: "TimeVaryingDetection_mrmr"
output: html_document
date: "2024-11-12"
---

**Here, I make a custom version of mrmr functons, aiming to get time-varying detection probability estimates. I updated the twostate.stan model to this end, and am now customizing mrmr's fit_model() function to use my updated stan model**

```{r}

library(cmdstanr)
library(mrmr)

```

**Write custom version of mrmr::fit_model()**

Change only line 53. Tell function to pull from my TVD_twostate.stan file

```{r}

# Custom fit function using the updated Stan model file
fit_model_custom <- function(
  data,
  seed = NULL,
  refresh = NULL,
  init = NULL,
  save_latent_dynamics = FALSE,
  output_dir = NULL,
  output_basename = NULL,
  sig_figs = NULL,
  chains = 4,
  parallel_chains = getOption("mc.cores", 1),
  chain_ids = seq_len(chains),
  threads_per_chain = NULL,
  opencl_ids = NULL,
  iter_warmup = 1000,
  iter_sampling = 1000,
  save_warmup = FALSE,
  thin = NULL,
  max_treedepth = NULL,
  adapt_engaged = TRUE,
  adapt_delta = NULL,
  step_size = NULL,
  metric = NULL,
  metric_file = NULL,
  inv_metric = NULL,
  init_buffer = NULL,
  term_buffer = NULL,
  window = NULL,
  fixed_param = FALSE,
  show_messages = TRUE,
  compile = TRUE,
  ...
) {
  # updated to use my custom stan model
  stan_file <- "updated_twostate.stan" # specify new path

  model <- cmdstan_model(stan_file, ...)

  m_fit <- model$sample(
    data = data$stan_d,
    seed = seed,
    refresh = refresh,
    init = init,
    save_latent_dynamics = save_latent_dynamics,
    output_dir = output_dir,
    output_basename = output_basename,
    sig_figs = sig_figs,
    chains = chains,
    parallel_chains = parallel_chains,
    chain_ids = chain_ids,
    threads_per_chain = threads_per_chain,
    opencl_ids = opencl_ids,
    iter_warmup = iter_warmup,
    iter_sampling = iter_sampling,
    save_warmup = save_warmup,
    thin = thin,
    max_treedepth = max_treedepth,
    adapt_engaged = adapt_engaged,
    adapt_delta = adapt_delta,
    step_size = step_size,
    metric = metric,
    metric_file = metric_file,
    inv_metric = inv_metric,
    init_buffer = init_buffer,
    term_buffer = term_buffer,
    window = window,
    fixed_param = fixed_param,
    show_messages = show_messages
  )

  # Load all the data and return the whole unserialized fit object
  m_fit$draws() # Do not specify variables or inc_warmup.
  try(m_fit$sampler_diagnostics(), silent = TRUE)
  try(m_fit$init(), silent = TRUE)
  try(m_fit$profiles(), silent = TRUE)

  list(m_fit = m_fit, data = data)
}

```

```{r}

# set paths for Trapdoor captures and surveys
captures <- read.csv("data/54188_capture.csv")  
surveys <- read.csv("data/54188_survey.csv")

data_TVD_model <- mrmr::clean_data(captures, surveys)
```

```{r}

# Fit the model using the custom function
fit_result <- fit_model_custom(data_TVD_model,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

modTVD_summary <- fit_result$m_fit$summary()
View(modTVD_summary)

write.csv(modTVD_summary, "Trapdoor_custom_summary")

```

**Now, run the classic mrmr model on the Trapdoor data to compare with my version**

```{r}

# Fit the model using the custom function
Trapdoor1_mrmr <- mrmr::fit_model(data_TVD_model,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

Trapdoor1_summary <- Trapdoor1_mrmr$m_fit$summary()
View(Trapdoor1_summary)

write.csv(Traodoor1_summary, "Trapdoor1_mrmr_summary.csv")

Trapdoor1_summary <- read.csv("Trapdoor1_mrmr_summary.csv")
View(trapdoor_summary)

```