### 2024-11-13 code/R/mrmr_mod1_mossy.Rmd

---
title: "Mossy Pond mrmr model 1"
output: GitHub_document
date: "`r Sys.Date()`"
author: "Your Name"
---

Load packages

```{r}

library(dplyr)
library(readr)
library(tibble)
library(rstan)
library(mrmr)

```

Load capture data and format 'captures' data frame

```{r}

# Load .csv of all captures from the Mossy Pond CMR study period
captures <- read.csv("Capture_Data.csv")

# Remove "00:00:00 EDT" from Date column for compatibility with mrmr 
captures$Date <- gsub("00:00:00 EDT","", captures$Date)

# Convert 'Date' variable to date object class 
captures$Date <- as.Date(captures$Date, format = "%a %b %d  %Y")

# Change variable names for compatibility with mrmr 
names(captures)[names(captures) == "Date"] <- "survey_date"
names(captures)[names(captures) == "PIT_TagCode"] <- "pit_tag_id"

# Change data types for compatibility with mrmr
captures$survey_date <- as.character(captures$survey_date)
captures$pit_tag_id <- as.character(captures$pit_tag_id)

# Remove the single dead frog observation (instead, use it to create a 'removals' data frame)
captures <- captures[captures$frog_state != "dead", ]

# ATTEMPT TO FIGURE THE PROBLEM OUT
# Include only a subset of columns from captures. This seems to have worked in the past. 
### captures <- unique(captures[ , c(2,3,5,7)])
  # when I create this subset, 5 observations disappear
  # This subset didn't work. Lets try getting rid of CA lake id column
captures <- unique(captures[ , c(3,5,7)])
  # another 2 observations disappeared
# I have not yet run this model

```

Load survey date data and format 'surveys' data frame

```{r}

# load surveys .csv I made externally (manually in excel...)
surveys <- read.csv("survey.csv")

# change data types for compatibility with mrmr
surveys$survey_date <- as.character(surveys$survey_date)
surveys$primary_period <- as.integer(surveys$primary_period)
surveys$secondary_period <- as.integer(surveys$secondary_period)

```

Load capture data and format 'removals' data frame

```{r}

removals <- read.csv("Capture_Data.csv")

removals <- removals[removals$frog_state == "dead", ]

# Remove "00:00:00 EDT" from Date column for compatibility with 'mrmr' 
removals$Date <- gsub("00:00:00 EDT","", removals$Date)

# Convert 'Date' variable to object class date
removals$Date <- as.Date(removals$Date, format = "%a %b %d  %Y")

# Change variable names for compatibility with 'mrmr' 
names(removals)[names(removals) == "Date"] <- "removal_date"
names(removals)[names(removals) == "PIT_TagCode"] <- "pit_tag_id"

removals$survey_date <- as.character(removals$removal_date)
removals$pit_tag_id <- as.character(removals$pit_tag_id)

```

Run the clean_data function

```{r}

# run the clean_data function
mrmr_mossy_mod1_data <- clean_data(captures, surveys)

```

fit the model

```{r}

# fit the model
mrmr_mossy_mod1 <- fit_model(mrmr_mossy_mod1_data,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

# save the model fit summary to an object
mod1_summary <- mrmr_mossy_mod1$m_fit$summary()
view(mod1_summary)

```

```{r}

plot_model(mrmr_mossy_mod1, "abundance")

```