### 2024-11-13 code/R/TimeVaryingDetection_mrmr.Rmd
plot and visually compare model output

---
title: "TimeVaryingDetection_mrmr"
output: html_document
date: "2024-11-12"
---

# Test my custom stan model with MLRG's public data from site 54188

    The modified stan model...

    -   allows for independent session-varying estimates of frog detection probabilities
    -   includes more parameter estimates in the generated quantities block
        -   p (detection probability, varying by session)
        -   phi_est (phi, or survival probability, estimated for the whole population between primary periods)

**Load packages**
```{r}
library(cmdstanr)
library(mrmr)
library(ggplot2)
library(ggpubr)
```

## Modify mrmr::fit_model()

-   fit_model_custom() is a copy of mrmr::fit_model(), changed only to call on a modified version of the underlying stan model.

```{r}

# Custom fit function using the updated Stan model file
fit_model_custom <- function(
  data,
  seed = NULL,
  refresh = NULL,
  init = NULL,
  save_latent_dynamics = FALSE,
  output_dir = NULL,
  output_basename = NULL,
  sig_figs = NULL,
  chains = 4,
  parallel_chains = getOption("mc.cores", 1),
  chain_ids = seq_len(chains),
  threads_per_chain = NULL,
  opencl_ids = NULL,
  iter_warmup = 1000,
  iter_sampling = 1000,
  save_warmup = FALSE,
  thin = NULL,
  max_treedepth = NULL,
  adapt_engaged = TRUE,
  adapt_delta = NULL,
  step_size = NULL,
  metric = NULL,
  metric_file = NULL,
  inv_metric = NULL,
  init_buffer = NULL,
  term_buffer = NULL,
  window = NULL,
  fixed_param = FALSE,
  show_messages = TRUE,
  compile = TRUE,
  ...
) {
  # updated to use my custom stan model
  stan_file <- "/Users/johnimperato/Desktop/Mossy-Pond/code/stan/JI_custom_twostate.stan" # specify local path

  model <- cmdstan_model(stan_file, ...)

  m_fit <- model$sample(
    data = data$stan_d,
    seed = seed,
    refresh = refresh,
    init = init,
    save_latent_dynamics = save_latent_dynamics,
    output_dir = output_dir,
    output_basename = output_basename,
    sig_figs = sig_figs,
    chains = chains,
    parallel_chains = parallel_chains,
    chain_ids = chain_ids,
    threads_per_chain = threads_per_chain,
    opencl_ids = opencl_ids,
    iter_warmup = iter_warmup,
    iter_sampling = iter_sampling,
    save_warmup = save_warmup,
    thin = thin,
    max_treedepth = max_treedepth,
    adapt_engaged = adapt_engaged,
    adapt_delta = adapt_delta,
    step_size = step_size,
    metric = metric,
    metric_file = metric_file,
    inv_metric = inv_metric,
    init_buffer = init_buffer,
    term_buffer = term_buffer,
    window = window,
    fixed_param = fixed_param,
    show_messages = show_messages
  )

  # Load all the data and return the whole unserialized fit object
  m_fit$draws() # Do not specify variables or inc_warmup.
  try(m_fit$sampler_diagnostics(), silent = TRUE)
  try(m_fit$init(), silent = TRUE)
  try(m_fit$profiles(), silent = TRUE)

  list(m_fit = m_fit, data = data)
}

```

## Run 54188 data through my custom model

**Load and clean data**
```{r}
# set paths for 54188 captures and surveys
captures <- read.csv("/Users/johnimperato/Desktop/Mossy-Pond/data/54188_capture.csv")  
surveys <- read.csv("/Users/johnimperato/Desktop/Mossy-Pond/data/54188_survey.csv")

# run mrmr::clean_data()
data_54188 <- mrmr::clean_data(captures, surveys)
```

```{r}
# Fit the custom model using the modified function
custom_mod_54188 <- fit_model_custom(data_custom_mod_54188,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

custom_mod_54188_summary <- custom_mod_54188$m_fit$summary()
View(custom_mod_54188_summary)

write.csv(custom_mod_54188_summary, "54188_custom_mod_summary")

```

**Plot abundance and recruitment**
```{r}

trapdoor_custom_abundance_plot <- mrmr::plot_model(custom_mod_54188, "abundance")
trapdoor_custom_recruitment_plot <- mrmr::plot_model(custom_mod_54188, "recruitment")

# arrange plots together
mrmr_custom_plots_54188 <- ggarrange(trapdoor_custom_abundance_plot, trapdoor_custom_recruitment_plot, ncol =1)

# add title
custom_annotated <- annotate_figure(mrmr_custom_plots_54188, top = text_grob("Time-Varying Detection mrmr", size = 14, face = "bold"))

```

## Run classic mrmr model on 54188 data

```{r}

# Fit the model using mrmr's built in function
classic_mod_54188 <- mrmr::fit_model(data_54188,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

classic_mod_54188_summary <- classic_mod_54188$m_fit$summary()
View(classic_mod_54188_summary)

write.csv(classic_mod_54188_summary, "54188_classic_mod_summary.csv")

plot_model(classic_mod_54188, "abundance")
plot_model(classic_mod_54188, "recruitment")

```

**Plot abundance and recruitment**
```{r}

trapdoor_classic_abundance_plot <- mrmr::plot_model(classic_mod_54188, "abundance")
trapdoor_classic_recruitment_plot <- mrmr::plot_model(classic_mod_54188, "recruitment")

mrmr_classic_plots_54188 <- ggarrange(trapdoor_classic_abundance_plot, trapdoor_classic_recruitment_plot, ncol =1)

# Add title
classic_annotated <- annotate_figure(mrmr_classic_plots_54188, top = text_grob("Classic mrmr", size = 14, face = "bold"))

```

## Compare Model Output

**Visual Comparison**
```{r}

abund_recruit_comparison_plot <- ggarrange(classic_annotated, custom_annotated, ncol=2)

```