### 2024-11-16 code/R/MossyPond_Custom_mrmr_TVD.Rmd
rename file

---
title: "Mossy_custom_mrmr"
output: html_document
date: "2024-11-13"
---

# Run my custom mrmr model with Mossy Pond data

```{r}
library(mrmr)
```

**Load and format data**

```{r}
getwd()
surveys <- read.csv("/Users/johnimperato/Desktop/Mossy-Pond/data/survey_MossyPond.csv")
captures <- read.csv("/Users/johnimperato/Documents/private_MP_data/captures_MP.csv")

mp_custom_data <- mrmr::clean_data(captures, surveys)

```

**Fit the custom model to MP data and summarize**

```{r}

# Fit the custom model using the modified function
custom_mod_MP <- fit_model_custom(mp_custom_data,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

custom_mod_mp_summary <- custom_mod_MP$m_fit$summary()
View(custom_mod_mp_summary)

write.csv(custom_mod_mp_summary, "/Users/johnimperato/Documents/mp_custom_mod_summary")

```

```{r}

abundance_MP_custom <- mrmr::plot_model(custom_mod_MP, "abundance")
recruitment_MP_custom <- mrmr::plot_model(custom_mod_MP, "recruitment")

```