### 2024-11-16 code/R/54188_ClassicCustomComparison_mrmr.Rmd
update formatting and clean up code

---
title: "TimeVaryingDetection_mrmr"
output: html_document
date: "2024-11-12"
---

# Model Comparison

**Classic mrmr model vs. custom mrmr model**

The custom model...

-   allows for independent session-varying estimates of frog detection probabilities
-   includes more parameter estimates in the generated quantities block
    -   p (detection probability, varying by session)
    -   phi_est (phi, or survival probability, estimated for the whole population between primary periods)

**Load packages**

```{r}
library(cmdstanr)
library(mrmr)
library(ggpubr)
```

## Run 54188 data through custom model

**Load and clean data**

```{r}
# set paths for 54188 captures and surveys
captures <- read.csv("/Users/johnimperato/Desktop/Mossy-Pond/data/54188_capture.csv")  
surveys <- read.csv("/Users/johnimperato/Desktop/Mossy-Pond/data/54188_survey.csv")

# run mrmr::clean_data()
data_54188 <- mrmr::clean_data(captures, surveys)
```

**Fit the custom model**

```{r}
# Fit the custom model using the modified function
custom_mod_54188 <- fit_model_custom(data_54188,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

# create and view model summary
custom_mod_54188_summary <- custom_mod_54188$m_fit$summary()
View(custom_mod_54188_summary)

# save results to .csv file
write.csv(custom_mod_54188_summary, "54188_custom_mod_summary")

```

**Plot abundance and recruitment**

```{r}

trapdoor_custom_abundance_plot <- mrmr::plot_model(custom_mod_54188, "abundance")
trapdoor_custom_recruitment_plot <- mrmr::plot_model(custom_mod_54188, "recruitment")

# arrange plots together
mrmr_custom_plots_54188 <- ggarrange(trapdoor_custom_abundance_plot, trapdoor_custom_recruitment_plot, ncol =1)

# add title
custom_annotated <- annotate_figure(mrmr_custom_plots_54188, top = text_grob("Time-Varying Detection mrmr", size = 14, face = "bold"))

```

## Run classic mrmr model on 54188 data

```{r}

# Fit the model using mrmr's built in function
classic_mod_54188 <- mrmr::fit_model(data_54188,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

classic_mod_54188_summary <- classic_mod_54188$m_fit$summary()
View(classic_mod_54188_summary)

write.csv(classic_mod_54188_summary, "54188_classic_mod_summary.csv")

plot_model(classic_mod_54188, "abundance")
plot_model(classic_mod_54188, "recruitment")

```

**Plot abundance and recruitment**

```{r}

trapdoor_classic_abundance_plot <- mrmr::plot_model(classic_mod_54188, "abundance")
trapdoor_classic_recruitment_plot <- mrmr::plot_model(classic_mod_54188, "recruitment")

mrmr_classic_plots_54188 <- ggarrange(trapdoor_classic_abundance_plot, trapdoor_classic_recruitment_plot, ncol =1)

# Add title
classic_annotated <- annotate_figure(mrmr_classic_plots_54188, top = text_grob("Classic mrmr", size = 14, face = "bold"))

```

## Compare Model Output

**Visual Comparison**

```{r}

abund_recruit_comparison_plot_54188 <- ggarrange(classic_annotated, custom_annotated, ncol=2)

```