### 2024-11-21 code/R/54188_ClassicCustomComparison_mrmr.Rmd
edit introduction

---
title: "TimeVaryingDetection_mrmr"
output: html_document
date: "2024-11-12"
---

# mrmr Modifications

# and

# Model Comparison

## Introduction

The R package mrmr serves as an interface to prepare data and run population models. It relies on a site-specific multi-state Bayesian open-population Jolly-Seber CMR model (hereafter referred to as the multi-state model) for the mathematical and computational part of model fitting. The multi-state model is written in the Stan programming language, which is designed for probabilistic modeling and Bayesian inference.

This report describes modifications to the multi-state model, each of which can be categorized as either 'cosmetic' or 'substantive.' Cosmetic changes are those that do not not affect how the model runs or the values of its parameter estimates. Substantive changes do affect how the model runs and the values of parameter estimates.

These modifications are embodied by a 'classic model' and a 'custom model.' The models are named relative to the original multi-state model published in the R package mrmr. The classic model includes only cosmetic changes relative to the original, and the custom model includes both cosmetic and substantive changes.

To understand the cosmetic and substantive nature of those changes, it is helpful to know that the multi-state Stan model is composed of seven **code blocks**: functions, data, transformed data, parameters, transformed parameters, model, and generated quantities. Changes made to the generated quantities block are cosmetic. Changes to any other block are substantive. This report details changes within the **generated quantities, parameters,** and **transformed parameters** blocks, as described in the next three sections.

### Generated Quantities Block

Both the classic and custom models include changes to the generated quantities block of mrmr's underlying Stan model. The generated quantities block is composed of post-hoc computations based on the model's parameters after it has finished fitting the data. Changes to this block are cosmetic and do not affect how the model runs or the estimated parameter values. The generated quantities blocks of both the custom and classic model are identical. Generated quantities include:

1.  N - total abundance, estimated for each primary period.

2.  B - number of adults recruited into the population, estimated for each primary period.

3.  p - detection probability, estimated for each session (secondary period). In the classic model, p is fixed so the estimate is identical for each session. In the custom model, p is estimated independently for each session (time-varying detection probabilities).

4.  phi_est - survival probability, estimated between primary periods for the whole population. phi_est[1] is the collective estimated survival probability between the first and second primary periods, phi_est[2] between the second and third, etc.\*

    -   \*I need to confirm my assumption that the model generates "boundary estimates" for phi (an estimate before the first primary period and after the last primary period). If I can't confirm this by myself, maybe I can ask Max.

5.  log_lik - log of the likelihood, calculated for each individual, real and pseudo. Likelihood is the probability of observing an individual's capture history given the model parameters. Pseudo-individuals don't actually have a capture histories, so their log_lik values are all identical. See the Compare Model Fit section of this report for more information about this quantity.

### Parameters Block

*vector[m_detect] beta_detect;* —\> *vector[Jtot] beta_detect;*

This substantive change is essential to achieving independent, time-varying estimates of detection probability in the custom model.

**beta_detect** is a vector of parameters that represent the logit-scale detection probability for survey sessions. The classic mrmr model defines beta_detect as a vector of size m_detect. m_detect refers to the number of detection covariates. The custom model defines beta_detect as a vector of size Jtot. Jtot refers to the total number of survey sessions. Each entry in the vector in the custom model corresponds to a specific survey session rather than a covariate, as in the classic model.

Meaning is assigned to beta_detect in the functions block via the transformed parameter logit_detect, within the forward_prob( ) function. No changes were made to this function.

### Transformed Parameters Block

(*logit_detect = X_detect \* beta_detect;*) —\> (*logit_detect = beta_detect;*)

In the classic model, logit_detect is derived from beta_detect using the design matrix of covariates X_detect. This ties detection probabilities to shared covariates. In the custom model, logit_detect is assigned directly to beta_detect, making each session's detection probability independent.

Changes to the parameters and transformed parameters blocks are substantive and aim for the most accurate estimates of N, B, p, and phi for the population as a whole. The classic version aims to model detection and survival by climate and disease-related covariates, which makes sense in the context of Knapp and Joseph (2018). If and when I want to introduce covariates into my analysis, I will need to revisit this section. (I'm not sure if it is possible to include both covariates in the design matrix and indepentent, session-specific detection.)

## The Data

The analysis in this report uses capture-mark-recapture (CMR) data for a population of endangered Sierra Nevada yellow-legged frogs (*Rana sierrae*) in the Sierra National Forest, California. This population is referred to by the numeric site id of its associated water body, 54188. The data set is publicly available from the Mountain Lakes Research Group.

## Analysis

This section describes the analysis performed to compare the classic and custom models. Both models were fit to the same CMR data, described in the previous section. Key comparisons include estimates of population abundance (N), recruitment (B), detection probability (p), and survival (phi). To evaluate model performance, I used the leave-one-out (LOO) cross-validation method, which assesses model fit while penalizing complexity.

The goal of this analysis is to determine whether the custom model, with independent, time-varying estimates of detection probability, provides a significantly better fit to the data compared to the classic model, which assumes detection probability is constant across sessions. Differences in model performance will be interpreted using the LOO expected log predictive density (elpd) values, along with associated uncertainty estimates.

Results and interpretations for each parameter and the overall model comparison are detailed in the following subsections.

**Load packages**

```{r}
library(cmdstanr)
library(mrmr)
library(ggpubr)
library(dplyr)
```

### Prepare data

Load data for population 54188 and format it with mrmr's clean_data( ) function.

```{r}
# set local paths for 54188 captures and surveys
captures <- read.csv("/Users/johnimperato/Desktop/Mossy-Pond/data/54188_capture.csv")  
surveys <- read.csv("/Users/johnimperato/Desktop/Mossy-Pond/data/54188_survey.csv")

# clean data
data_54188 <- mrmr::clean_data(captures, surveys)
```

### Fit the Custom Model

The fit_model_custom( ) function is identical to mrmr's built in fit_model( ) function, except it draws on the custom .stan file on my local machine, rather than the standard mrmr .stan file. The custom .stan file includes the previously described changes to the parameters, transformed parameters, and generated quantities blocks.

```{r}
# load the modified fit_model function
source("custom_fit_model_function.R")

# Fit the custom model using the modified function
custom_mod_54188 <- fit_model_custom(data_54188,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

# create and view model summary
custom_mod_54188_summary <- custom_mod_54188$m_fit$summary()
print(custom_mod_54188_summary)

# save results to .csv file
write.csv(custom_mod_54188_summary, "54188_custom_mod_summary")

```

**Plot abundance and recruitment**

...with mrmr's built-in plotting functionality.

```{r}
trapdoor_custom_abundance_plot <- mrmr::plot_model(custom_mod_54188, "abundance")
trapdoor_custom_recruitment_plot <- mrmr::plot_model(custom_mod_54188, "recruitment")

# arrange plots 
mrmr_custom_plots_54188 <- ggarrange(trapdoor_custom_abundance_plot, trapdoor_custom_recruitment_plot, ncol =1)

# add title
custom_annotated <- annotate_figure(mrmr_custom_plots_54188, top = text_grob("Time-Varying Detection mrmr", size = 14, face = "bold"))
```

### Fit the classic model

The fit_model_classic( ) function is identical to mrmr's built in fit_model( ) function, except it draws on a modified .stan file on my local machine, rather than the standard mrmr .stan file. The modified .stan file includes the previously described changes to the generated quantities block.

```{r}
source("classic_fit_model_function.R")

# Fit the model using mrmr's built in function
classic_mod_54188 <- fit_model_classic(data_54188,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

# create and view summary
classic_mod_54188_summary <- classic_mod_54188$m_fit$summary()
print(classic_mod_54188_summary)

# save summary to .csv file
write.csv(classic_mod_54188_summary, "54188_classic_mod_summary.csv")

```

**Plot abundance and recruitment**

...with mrmr's built-in plotting functionality.

```{r}
trapdoor_classic_abundance_plot <- mrmr::plot_model(classic_mod_54188, "abundance")
trapdoor_classic_recruitment_plot <- mrmr::plot_model(classic_mod_54188, "recruitment")

# arrange plots
mrmr_classic_plots_54188 <- ggarrange(trapdoor_classic_abundance_plot, trapdoor_classic_recruitment_plot, ncol =1)

# add title
classic_annotated <- annotate_figure(mrmr_classic_plots_54188, top = text_grob("Classic mrmr", size = 14, face = "bold"))

```

### Model Comparison

#### Generated Quantities

**Create a table comparing the mean generated quantity values for each model**

This table shows **only the mean** estimated value for each generated quantity. It is meant to serve as a quick reference for comparing model output. See the individual model summaries for complete generated quantity summaries.

```{r}
# pull quantities from model fit summary
N_classic <- classic_mod_54188_summary[4216:4220, 1:2]
N_custom <- custom_mod_54188_summary[4229:4233, 1:2]
B_classic <- classic_mod_54188_summary[4221:4225, 1:2]
B_custom <- custom_mod_54188_summary[4234:4238, 1:2]
Phi_classic <- classic_mod_54188_summary[4240:4243, 1:2]
Phi_custom <- custom_mod_54188_summary[4253:4256, 1:2]
p_classic <- classic_mod_54188_summary[4226:4239, 1:2]
p_custom <- custom_mod_54188_summary[4239:4252, 1:2]

GenQuant_comparison <- bind_rows(
  left_join(N_classic, N_custom, by = "variable", suffix = c("_classic", "_custom")) %>% mutate(parameter = "abundance"),
  left_join(B_classic, B_custom, by = "variable", suffix = c("_classic", "_custom")) %>% mutate(parameter = "recruitment"),
  left_join(Phi_classic, Phi_custom, by = "variable", suffix = c("_classic", "_custom")) %>% mutate(parameter = "survival"),
  left_join(p_classic, p_custom, by = "variable", suffix = c("_classic", "_custom")) %>% mutate(parameter = "detection"),
) %>% 
  select(variable, parameter, everything())

print(GenQuant_comparison)

```

**Visually compare abundance and recruitment estimates between models**

Note that axes are not standardized.

```{r}
# visual comparison of abundance and recruitment plots
abund_recruit_comparison_plot_54188 <- ggarrange(classic_annotated, custom_annotated, ncol=2)
print(abund_recruit_comparison_plot_54188)
```

#### Model Fit

Use leave-one-out (LOO) cross-validation, a standard method to calculate the predictive accuracy of Bayesian models, to compare how well the classic and custom models fit the Trapdoor data. The cmdstanr package has built in LOO capabilities. Briefly,

-   LOO cross-validation uses log likelihood values, which now appear in the model summary as generated quantities, to evaluate how well the model predicts each data point when that point is left out of the fitting process. Likelihood, calculated at the individual level, is the probability of observing the data (in this case, an individual frog's capture history) given the model parameters.

The loo( ) function generates an expected log predictive density (elpd) value for each model. The loo_compare( ) function ranks models based on their elpd values, with the best fit listed first, and calculates the difference in those values. A difference of more than 2x the standard error is generally considered significant in a LOO comparison.

```{r}
loo_classic <- classic_mod_54188$m_fit$loo()

loo_custom <- custom_mod_54188$m_fit$loo()

loo_compare <- loo::loo_compare(loo_classic, loo_custom)

print(loo_compare)

# extract elpd_diff and se_diff for model1
elpd_diff_model1 <- loo_compare["model1", "elpd_diff"]
se_diff_model1 <- loo_compare["model1", "se_diff"]

# calculate significance ratio
result <- elpd_diff_model1 / se_diff_model1

# interpret
if (abs(result) > 2) {
  cat("Result of elpd_diff / se_diff for model1:", result, "\n")
  cat("The custom model fits the data significantly better than the classic model.\n")
} else {
  cat("Result of elpd_diff / se_diff for model1:", result, "\n")
  cat("The custom model does not fit the data significantly better than the classic model.\n")
}

```