### 2024-11-22 code/R/MossyPond_Classic_mrmr.Rmd
update fit_model_classic function and change object names

---
title: "Mossy Pond mrmr model 1"
output: GitHub_document
date: "`r Sys.Date()`"
author: "Your Name"
---

```{r}
library(dplyr)
library(readr)
library(tibble)
library(rstan)
library(mrmr)
library(ggpubr)
```

```{r}

# load data
captures <- read.csv('/Users/johnimperato/Desktop/Mossy-Pond_companion/MP_captures.csv')
surveys <- read.csv('/Users/johnimperato/Desktop/Mossy-Pond/data/survey_MossyPond.csv')

# clean data 
mp_data <- clean_data(captures, surveys)

# load fit_model_classic function
source("classic_fit_model_function.R")

# fit classic mrmr model
classic_mod_mp <- fit_model_classic(mp_data,
                                chains = 3, 
                                iter_sampling = 1000, 
                                iter_warmup = 400)

# save model summary to an object
classic_mod_mp_summary <- classic_mod_mp$m_fit$summary()
View(classic_mod_mp_summary)

write.csv(classic_mod_mp_summary, "/Users/johnimperato/Desktop/Mossy-Pond_companion/mp_classic_mod_sumary.csv")

```

```{r}

abundance_MP_classic <- plot_model(mrmr_mossy_mod1, "abundance")
recruitment_MP_classic <- plot_model(mrmr_mossy_mod1, "recruitment")

MP_classic_plots <- ggarrange(abundance_MP_classic, recruitment_MP_classic, ncol=1)
MP_classic_plots

```