### 2024-12-05 code/R/WinterSeverity_analysis_MossyPond.Rmd
Initial commit of Winter Severity - Vital Rates analysis for Mossy Pond

---
title: "Yearly 1 April SWE and Vital Rates"
output: html_document
date: "2024-12-04"
---

# Analysis of Winter Severity and *Rana sierrae* Population Characteristics

**Winter severity measure**: percent of historical average 1 April SWE in the Mossy Pond study area.

-   snow data taken from 10 observation stations between 5.1 and 19.5 km from the MPSA

## Yearly % of historical average 1 April SWE in the Mossy Pond study area

```{r}
library(ggplot2)
```

```{r}

# create winter severity data frame
avg_swe <- read.csv('/Users/johnimperato/Desktop/Mossy-Pond/data/Yearly_WinterSeverity_MPSA.csv')

avg_swe <- avg_swe %>%
  rename(Winter_Severity = Average.Pixel.Value) %>%
  mutate(Winter = c("2011-2012", "2012-2013", "2013-2014", "2014-2015", "2015-2016", "2016-2017", "2017-2018"))

# create data frame for winter severity one year prior
avg_swe_year_prior <- avg_swe %>%
                        mutate(Year = Year + 1)

```

## Population Vital Rates/Numbers and Growth Rate

**N, B, and phi**

```{r}

model_summary <- read.csv('/Users/johnimperato/Desktop/Mossy-Pond_companion/mp_custom_mod_sumary.csv')

# extract N values from model summary
N <- model_summary[33659:33674, ]

# calculate yearly average N
N_yearly_avg <- N %>%
  # Group rows and calculate the mean of numeric columns
  mutate(group = c(rep(1:4, each = 3), rep(5, 4))) %>%
  group_by(group) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE)) %>%
  ungroup() %>%
  select(-group) %>%
  # Add a new Year column
  mutate(Year = c(2014:2018)) %>%
  # Rename 'mean' to 'Abundance' and reorder columns
  rename(Mean_Abundance = mean) %>%
  select(Year, Mean_Abundance)

# extract B values from model summary
B <- model_summary[c(33675:33690), ]

# calculate yearly total B
B_yearly_avg <- B %>%
  # Group rows and calculate the mean of numeric columns
  mutate(group = c(rep(1:4, each = 3), rep(5, 4))) %>%
  group_by(group) %>%
  summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
  ungroup() %>%
  select(-group) %>%
  # Add a new Year column
  mutate(Year = c(2014:2018)) %>%
  # Rename 'mean' to 'Recruited_Individuals' and reorder columns
  rename(Recruited_Individuals = mean) %>%
  select(Year, Recruited_Individuals)

# extract overwinter survival values from model summary
phi <- B <- model_summary[c(33738:33752), ]

# extract overwinter survival
phi_overwinter <- phi %>%
  slice(c(3, 6, 9, 12)) %>%
  # Add a new Year column
  mutate(Winter = c("2014-2015", "2015-2016", "2016-2017", "2017-2018")) %>%
  # Rename 'mean' to 'Recruited_Individuals' and reorder columns
  rename(Overwinter_Survival_Probability = mean) %>%
  select(Winter, Overwinter_Survival_Probability)

```

**Growth rate**

```{r}

```

## Abundance-Winter severity Comaprison

### Winter 1 year ago

```{r}

# combine data fames
comparison_df_1 <- N_yearly_avg %>%
  left_join(avg_swe_year_prior, by = "Year")

# plot
abundance_WinterSeverity_1yr <- ggplot(comparison_df_1, aes(x = Year)) +
  # frog abundance line
  geom_line(aes(y = Mean_Abundance, color = "Frog Abundance"), size = 1) +
  geom_point(aes(y = Mean_Abundance, color = "Frog Abundance"), size = 2) +
  # winter severity line
  geom_line(aes(y = Winter_Severity, color = "Winter Severity One Year Prior\n(% of historical avg 1 April SWE)"), linetype = "dashed", size = 1) +
  geom_point(aes(y = Winter_Severity, color = "Winter Severity One Year Prior\n(% of historical avg 1 April SWE)"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Frog Abundance" = "blue", "Winter Severity One Year Prior\n(% of historical avg 1 April SWE)" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Frog Abundance and Winter Severity One Year Prior",
    x = "Year",
    y = "Value"
  )

# correlation
correlation1 <- cor(comparison_df_1$Mean_Abundance, comparison_df_1$Winter_Severity)
cor_test1 <- cor.test(comparison_df_1$Mean_Abundance, comparison_df_1$Winter_Severity)

correlation_plot1 <- ggplot(data = comparison_df_1, aes(x = Winter_Severity, y = Mean_Abundance)) +
  geom_point() +
  geom_smooth(method = "lm" ,se = FALSE) +
  annotate("text", x = 75, y = 185, label = paste0("Correlation: ", round(correlation1, 2)), size = 5, color = "red") +
  labs(
    title = "Correlation between Winter Severity and Abundance",
    x = "Winter Severity (% of Historical Average 1 April SWE)",
    y = "Yearly Mean Abundance"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

## Recruitment - Winter Severity Comparison

### Winter 1 year ago

```{r}

# combine data fames
comparison_df_2 <- B_yearly_avg %>%
  left_join(avg_swe_year_prior, by = "Year")

# plot
WinterSeverity_Recruitment <- ggplot(comparison_df_2, aes(x = Year)) +
  # frog abundance line
  geom_line(aes(y = Recruited_Individuals, color = "Individuals Recruited"), size = 1) +
  geom_point(aes(y = Recruited_Individuals, color = "Individuals Recruited"), size = 2) +
  # winter severity line
  geom_line(aes(y = Winter_Severity, color = "Winter Severity One Year Prior\n(% of historical avg 1 April SWE)"), linetype = "dashed", size = 1) +
  geom_point(aes(y = Winter_Severity, color = "Winter Severity One Year Prior\n(% of historical avg 1 April SWE)"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Individuals Recruited" = "blue", "Winter Severity One Year Prior\n(% of historical avg 1 April SWE)" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Recruitment and Winter Severity One Year Prior",
    x = "Year",
    y = "Value"
  )

# correlation
correlation2 <- cor(comparison_df_2$Winter_Severity, comparison_df_2$Recruited_Individuals)
cor_test2 <- cor.test(comparison_df_1$Winter_Severity, comparison_df_2$Recruited_Individuals)

correlation_plot2 <- ggplot(data = comparison_df_2, aes(x = Winter_Severity, y = Recruited_Individuals)) +
  geom_point() +
  geom_smooth(method = "lm" ,se = FALSE) +
  annotate("text", x = 75, y = 185, label = paste0("Correlation: ", round(correlation2, 2)), size = 5, color = "red") +
  labs(
    title = "Correlation between Winter Severity and Recruitment",
    x = "Winter Severity (% of Historical Average 1 April SWE)",
    y = "Total Yearly Recruitment (Number of individuals)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

## Survival Probability - Winter Severity Comparison

### Same Winter

```{r}

# combine data fames
comparison_df_3 <- phi_overwinter %>%
  left_join(avg_swe, by = "Winter")

# Ensure Winter is a factor in the correct order
comparison_df_3 <- comparison_df_3 %>%
  mutate(Winter = factor(Winter, levels = c("2014-2015", "2015-2016", "2016-2017", "2017-2018"))) %>%
  mutate(Survival_Percentage = Overwinter_Survival_Probability * 100)

# Plot
ggplot(comparison_df_3, aes(x = Winter)) +
  # survival line
  geom_line(aes(y = Survival_Percentage, color = "Survival Percentage"), size = 1, group = 1) +
  geom_point(aes(y = Survival_Percentage, color = "Survival Percentage"), size = 2) +
  # winter severity line
  geom_line(aes(y = Winter_Severity, color = "Winter Severity"), linetype = "dashed", size = 1, group = 2) +
  geom_point(aes(y = Winter_Severity, color = "Winter Severity"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Survival Percentage" = "blue", "Winter Severity" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Survival and Winter Severity",
    x = "Winter",
    y = "Value"
  )

```