### 2025-01-17 code/R/WinterSeverity_analysis_MossyPond.Rmd
add code for analaysis with 3-yr avg winters

---
title: "Yearly 1 April SWE and Vital Rates"
output: html_document
date: "2024-12-04"
---

# Analysis of Winter Severity and *Rana sierrae* Population Characteristics

**Winter severity measure**: percent of historical average 1 April SWE in the Mossy Pond study area.

-   1 April SWE data interpolated from 10 observation stations between 5.1 and 19.5 km from the MPSA
-   Data downloaded from: <https://cdec.water.ca.gov/dynamicapp/wsSensorData>

## Yearly % of historical average 1 April SWE in the Mossy Pond study area

```{r}
library(ggplot2)
library(tidyverse)
library(ggpubr)
```

```{r}

# create winter severity data frame
avg_swe <- read.csv('/Users/johnimperato/Desktop/Mossy-Pond/data/Yearly_Winter_Severity_2011_2018_MPSA.csv')

avg_swe <- avg_swe %>%
  rename(Winter_Severity = Average.Pixel.Value) %>%
  mutate(Winter = c("2010-2011", "2011-2012", "2012-2013", "2013-2014", "2014-2015", "2015-2016", "2016-2017", "2017-2018"))

# create data frame for winter severity one year prior
avg_swe_year_prior <- avg_swe %>%
                        mutate(Year = Year + 1)

# create data frame for winter severity two years prior
avg_swe_2year_prior <- avg_swe %>%
                        mutate(Year = Year + 2)

# three-year rolling average 
avg_swe_3yrAvg <- avg_swe %>% 
  mutate(
    swe_3yrAvg = (
      Winter_Severity + 
      lag(Winter_Severity, 1) +
      lag(Winter_Severity, 2)
      ) / 3
  )

# three-year rolling average with one-year lag
avg_swe_3yrAvg_1yrLag <- avg_swe %>% 
  mutate(
    swe_3yrAvg_1yrLag = (
      lag(Winter_Severity, 1) +
      lag(Winter_Severity, 2) +
      lag(Winter_Severity, 3)
      ) / 3
  )

```

## Population Vital Rates/Numbers and Growth Rate

**N, B, and phi**

```{r}

model_summary <- read.csv('/Users/johnimperato/Desktop/Mossy-Pond_companion/mp_custom_mod_sumary.csv')

# extract N values from model summary
N <- model_summary[33659:33674, ]

# calculate yearly average N
N_yearly_avg <- N %>%
  # Group rows and calculate the mean of numeric columns
  mutate(group = c(rep(1:4, each = 3), rep(5, 4))) %>%
  group_by(group) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE)) %>%
  ungroup() %>%
  select(-group) %>%
  # Add a new Year column
  mutate(Year = c(2014:2018)) %>%
  # Rename 'mean' to 'Abundance' and reorder columns
  rename(Mean_Abundance = mean) %>%
  select(Year, Mean_Abundance)

# calculate yearly median N
N_yearly_median <- N %>%
  # Group rows and calculate the mean of numeric columns
  mutate(group = c(rep(1:4, each = 3), rep(5, 4))) %>%
  group_by(group) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE)) %>%
  ungroup() %>%
  select(-group) %>%
  # Add a new Year column
  mutate(Year = c(2014:2018)) %>%
  # Rename 'mean' to 'Abundance' and reorder columns
  rename(Mean_Abundance = median) %>%
  select(Year, Mean_Abundance)

# extract B values from model summary
B <- model_summary[c(33675:33690), ]

# calculate yearly total B
B_yearly_avg <- B %>%
  # Group rows and calculate the mean of numeric columns
  mutate(group = c(rep(1:4, each = 3), rep(5, 4))) %>%
  group_by(group) %>%
  summarise(across(where(is.numeric), sum, na.rm = TRUE)) %>%
  ungroup() %>%
  select(-group) %>%
  # Add a new Year column
  mutate(Year = c(2014:2018)) %>%
  # Rename 'mean' to 'Recruited_Individuals' and reorder columns
  rename(Recruited_Individuals = mean) %>%
  select(Year, Recruited_Individuals)

# extract overwinter survival values from model summary
phi <- B <- model_summary[c(33738:33752), ]

# extract overwinter survival
phi_overwinter <- phi %>%
  slice(c(3, 6, 9, 12)) %>%
  # Add a new Year column
  mutate(Winter = c("2014-2015", "2015-2016", "2016-2017", "2017-2018")) %>%
  # Rename 'mean' to 'Recruited_Individuals' and reorder columns
  rename(Overwinter_Survival_Probability = mean) %>%
  select(Winter, Overwinter_Survival_Probability)

print(N)
print(B)

print(B_yearly_avg)
print(N_yearly_avg)
print(phi_overwinter)

```

## Abundance-Winter severity Comaprison

### Same Year

```{r}

# combine data fames
comparison_df_N0 <- N_yearly_avg %>%
  left_join(avg_swe, by = "Year")

# plot
WinterSeverity_N0 <- ggplot(comparison_df_N0, aes(x = Year)) +
  # frog abundance line
  geom_line(aes(y = Mean_Abundance, color = "Frog Abundance"), size = 1) +
  geom_point(aes(y = Mean_Abundance, color = "Frog Abundance"), size = 2) +
  # winter severity line
  geom_line(aes(y = Winter_Severity, color = "Winter Severity Same Year\n(% of historical avg 1 April SWE)"), linetype = "dashed", size = 1) +
  geom_point(aes(y = Winter_Severity, color = "Winter Severity Same Year\n(% of historical avg 1 April SWE)"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Frog Abundance" = "blue", "Winter Severity Same Year\n(% of historical avg 1 April SWE)" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Frog Abundance and Winter Severity Same Year",
    x = "Year",
    y = "Value"
  )

# correlation
correlation_N0 <- cor(comparison_df_N0$Mean_Abundance, comparison_df_N0$Winter_Severity)
cor_test_N0 <- cor.test(comparison_df_N0$Mean_Abundance, comparison_df_N0$Winter_Severity)

correlation_plot_N0 <- ggplot(data = comparison_df_S_N, aes(x = Winter_Severity, y = Mean_Abundance)) +
  geom_point() +
  geom_smooth(method = "lm" ,se = FALSE) +
  annotate("text", x = 75, y = 185, label = paste0("Correlation: ", round(correlation_N0, 2)), size = 5, color = "red") +
  labs(
    title = "Correlation between Same-Year Winter Severity and Abundance",
    x = "Winter Severity (% of Historical Average 1 April SWE)",
    y = "Yearly Mean Abundance"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

######################## Print
print(WinterSeverity_N0)
print(correlation_plot_N0)

```

### One year ago

```{r}

# combine data fames
comparison_df_N1 <- N_yearly_avg %>%
  left_join(avg_swe_year_prior, by = "Year")

###################### plot
WinterSeverity_N1 <- ggplot(comparison_df_N1, aes(x = Year)) +
  # frog abundance line
  geom_line(aes(y = Mean_Abundance, color = "Frog Abundance"), size = 1) +
  geom_point(aes(y = Mean_Abundance, color = "Frog Abundance"), size = 2) +
  # winter severity line
  geom_line(aes(y = Winter_Severity, color = "Winter Severity One Year Prior\n(% of historical avg 1 April SWE)"), linetype = "dashed", size = 1) +
  geom_point(aes(y = Winter_Severity, color = "Winter Severity One Year Prior\n(% of historical avg 1 April SWE)"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Frog Abundance" = "blue", "Winter Severity One Year Prior\n(% of historical avg 1 April SWE)" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Frog Abundance and Winter Severity One Year Prior",
    x = "Year",
    y = "Value"
  ) +
  theme(legend.position = c(0.8, 0.2))

###################### correlation
correlation_N1 <- cor(comparison_df_N1$Mean_Abundance, comparison_df_N1$Winter_Severity)
cor_test_1_N <- cor.test(comparison_df_N1$Mean_Abundance, comparison_df_N1$Winter_Severity)

correlation_plot_N1 <- ggplot(data = comparison_df_N1, aes(x = Winter_Severity, y = Mean_Abundance)) +
  geom_point() +
  geom_smooth(method = "lm" ,se = FALSE) +
  annotate("text", x = 75, y = 185, label = paste0("Correlation: ", round(correlation_N1, 2)), size = 5, color = "red") +
  labs(
    title = "Correlation between Winter Severity and Abundance",
    x = "Winter Severity (% of Historical Average 1 April SWE)",
    y = "Yearly Mean Abundance"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

###################### print
print(WinterSeverity_N1)
print(correlation_plot_N1)

ggsave("WinterSeverityN1.png", plot = WinterSeverity_N1, width = 7, height = 4, bg = "white")
ggsave("CorrelationN1.png", plot = correlation_plot_N1, width = 7, height = 4, bg = "white")

N1_arrange <- ggarrange(WinterSeverity_N1, correlation_plot_N1)
ggsave("test12345.png", plot = N1_arrange, width = 10, bg = "white")

```

### Two Years Ago

```{r}

# combine data fames
comparison_df_2_N <- N_yearly_avg %>%
  left_join(avg_swe_2year_prior, by = "Year")

# plot
abundance_WinterSeverity_2yr <- ggplot(comparison_df_2_N, aes(x = Year)) +
  # frog abundance line
  geom_line(aes(y = Mean_Abundance, color = "Frog Abundance"), size = 1) +
  geom_point(aes(y = Mean_Abundance, color = "Frog Abundance"), size = 2) +
  # winter severity line
  geom_line(aes(y = Winter_Severity, color = "Winter Severity Two Years Prior\n(% of historical avg 1 April SWE)"), linetype = "dashed", size = 1) +
  geom_point(aes(y = Winter_Severity, color = "Winter Severity Two Years Prior\n(% of historical avg 1 April SWE)"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Frog Abundance" = "blue", "Winter Severity Two Years Prior\n(% of historical avg 1 April SWE)" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Frog Abundance and Winter Severity Two Years Prior",
    x = "Year",
    y = "Value"
  )

# correlation
correlation_2_N <- cor(comparison_df_2_N$Mean_Abundance, comparison_df_2_N$Winter_Severity)
cor_test_2_N <- cor.test(comparison_df_2_N$Mean_Abundance, comparison_df_2_N$Winter_Severity)

correlation_plot_2_N <- ggplot(data = comparison_df_2_N, aes(x = Winter_Severity, y = Mean_Abundance)) +
  geom_point() +
  geom_smooth(method = "lm" ,se = FALSE) +
  annotate("text", x = 50, y = 185, label = paste0("Correlation: ", round(correlation_2_N, 2)), size = 5, color = "red") +
  labs(
    title = "Correlation between Winter Severity Two Years Prior and Abundance",
    x = "Winter Severity (% of Historical Average 1 April SWE)",
    y = "Yearly Mean Abundance"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```

### Three-year rolling average

```{r}

# combine data fames
comparison_df_N3 <- N_yearly_avg %>%
  left_join(avg_swe_3yrAvg, by = "Year")

# plot
WinterSeverity_N3 <- ggplot(comparison_df_N3, aes(x = Year)) +
  # frog abundance line
  geom_line(aes(y = Mean_Abundance, color = "Frog Abundance"), size = 1) +
  geom_point(aes(y = Mean_Abundance, color = "Frog Abundance"), size = 2) +
  # winter severity line
  geom_line(aes(y = swe_3yrAvg, color = "Winter Severity 3 yr avg\n(% of historical avg 1 April SWE)"), linetype = "dashed", size = 1) +
  geom_point(aes(y = swe_3yrAvg, color = "Winter Severity 3 yr avg\n(% of historical avg 1 April SWE)"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Frog Abundance" = "blue", "Winter Severity 3 yr avg\n(% of historical avg 1 April SWE)" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Frog Abundance and Winter Severity Same Year",
    x = "Year",
    y = "Value"
  )

# correlation
correlation_N3 <- cor(comparison_df_N3$Mean_Abundance, comparison_df_N3$swe_3yrAvg)
cor_test_N3 <- cor.test(comparison_df_N0$Mean_Abundance, comparison_df_N3$swe_3yrAvg)

correlation_plot_N3 <- ggplot(data = comparison_df_N3, aes(x = swe_3yrAvg, y = Mean_Abundance)) +
  geom_point() +
  geom_smooth(method = "lm" ,se = FALSE) +
  annotate("text", x = 75, y = 185, label = paste0("Correlation: ", round(correlation_N3, 2)), size = 5, color = "red") +
  labs(
    title = "Correlation between 3-yr avg Winter Severity and Abundance",
    x = "Winter Severity (% of Historical Average 1 April SWE)",
    y = "Yearly Mean Abundance"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

######################## Print
print(WinterSeverity_N3)
print(correlation_plot_N3)

```
### Three-year rolling average w/ 1-yr lag

```{r}

# combine data fames
comparison_df_N3.1 <- N_yearly_avg %>%
  left_join(avg_swe_3yrAvg_1yrLag, by = "Year")

# plot
WinterSeverity_N3.1 <- ggplot(comparison_df_N3.1, aes(x = Year)) +
  # frog abundance line
  geom_line(aes(y = Mean_Abundance, color = "Frog Abundance"), size = 1) +
  geom_point(aes(y = Mean_Abundance, color = "Frog Abundance"), size = 2) +
  # winter severity line
  geom_line(aes(y = swe_3yrAvg_1yrLag, color = "Winter Severity 3 yr avg 1 yr lag\n(% of historical avg 1 April SWE)"), linetype = "dashed", size = 1) +
  geom_point(aes(y = swe_3yrAvg_1yrLag, color = "Winter Severity 3 yr avg 1 yr lag\n(% of historical avg 1 April SWE)"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Frog Abundance" = "blue", "Winter Severity 3 yr avg 1 yr lag\n(% of historical avg 1 April SWE)" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Frog Abundance and Winter Severity Same Year",
    x = "Year",
    y = "Value"
  )

# correlation
correlation_N3.1 <- cor(comparison_df_N3.1$Mean_Abundance, comparison_df_N3.1$swe_3yrAvg_1yrLag)
cor_test_N3.1 <- cor.test(comparison_df_N0$Mean_Abundance, comparison_df_N3.1$swe_3yrAvg_1yrLag)

correlation_plot_N3.1 <- ggplot(data = comparison_df_N3.1, aes(x = swe_3yrAvg_1yrLag, y = Mean_Abundance)) +
  geom_point() +
  geom_smooth(method = "lm" ,se = FALSE) +
  annotate("text", x = 75, y = 185, label = paste0("Correlation: ", round(correlation_N3.1, 2)), size = 5, color = "red") +
  labs(
    title = "Correlation between 3-yr avg Winter Severity with 1 yr lag and Abundance",
    x = "Winter Severity (% of Historical Average 1 April SWE)",
    y = "Yearly Mean Abundance"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

######################## Print
print(WinterSeverity_N3.1)
print(correlation_plot_N3.1)

```

## Recruitment - Winter Severity Comparison

### Winter Same Year

```{r}

# combine data fames
comparison_df_S_B <- B_yearly_avg %>%
  left_join(avg_swe, by = "Year")

# plot
WinterSeverity_S_B <- ggplot(comparison_df_S_B, aes(x = Year)) +
  # frog abundance line
  geom_line(aes(y = Recruited_Individuals, color = "Individuals Recruited"), size = 1) +
  geom_point(aes(y = Recruited_Individuals, color = "Individuals Recruited"), size = 2) +
  # winter severity line
  geom_line(aes(y = Winter_Severity, color = "Winter Severity Same Year"), linetype = "dashed", size = 1) +
  geom_point(aes(y = Winter_Severity, color = "Winter Severity Same Year"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Individuals Recruited" = "blue", "Winter Severity Same Year" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Recruitment and Same-Year Winter Severity",
    x = "Year",
    y = "Value"
  )

# correlation
correlation_S_B <- cor(comparison_df_S_B$Winter_Severity, comparison_df_S_B$Recruited_Individuals)
cor_test_S_B <- cor.test(comparison_df_S_B$Winter_Severity, comparison_df_S_B$Recruited_Individuals)

correlation_plot_S_B <- ggplot(data = comparison_df_S_B, aes(x = Winter_Severity, y = Recruited_Individuals)) +
  geom_point() +
  geom_smooth(method = "lm" ,se = FALSE) +
  annotate("text", x = 75, y = 185, label = paste0("Correlation: ", round(correlation_S_B, 2)), size = 5, color = "red") +
  labs(
    title = "Correlation between Same-Year Winter Severity and Recruitment",
    x = "Winter Severity (% of Historical Average 1 April SWE)",
    y = "Total Yearly Recruitment (Number of individuals)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

print(WinterSeverity_S_B)
print(correlation_plot_S_B)

```

  ### Winter 1 year ago

```{r}

# combine data fames
comparison_df_B1 <- B_yearly_avg %>%
  left_join(avg_swe_year_prior, by = "Year")

# plot
WinterSeverity_B1 <- ggplot(comparison_df_B1, aes(x = Year)) +
  # frog abundance line
  geom_line(aes(y = Recruited_Individuals, color = "Individuals Recruited"), size = 1) +
  geom_point(aes(y = Recruited_Individuals, color = "Individuals Recruited"), size = 2) +
  # winter severity line
  geom_line(aes(y = Winter_Severity, color = "Winter Severity One Year Prior\n(% of historical avg 1 April SWE)"), linetype = "dashed", size = 1) +
  geom_point(aes(y = Winter_Severity, color = "Winter Severity One Year Prior\n(% of historical avg 1 April SWE)"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Individuals Recruited" = "blue", "Winter Severity One Year Prior\n(% of historical avg 1 April SWE)" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Recruitment and Winter Severity One Year Prior",
    x = "Year",
    y = "Value"
  ) +
  theme(legend.position = c(0.8, 0.2))

# correlation
correlation_B1 <- cor(comparison_df_B1$Winter_Severity, comparison_df_B1$Recruited_Individuals)
cor_test_B1 <- cor.test(comparison_df_B1$Winter_Severity, comparison_df_B1$Recruited_Individuals)

correlation_plot_B1 <- ggplot(data = comparison_df_B1, aes(x = Winter_Severity, y = Recruited_Individuals)) +
  geom_point() +
  geom_smooth(method = "lm" ,se = FALSE) +
  annotate("text", x = 75, y = 185, label = paste0("Correlation: ", round(correlation_B1, 2)), size = 5, color = "red") +
  labs(
    title = "Correlation between Winter Severity and Recruitment",
    x = "Winter Severity (% of Historical Average 1 April SWE)",
    y = "Total Yearly Recruitment (Number of individuals)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        )

print(WinterSeverity_B1)
print(correlation_plot_B1)

ggsave("WinterSeverityB1.png", plot = WinterSeverity_B1, bg = "white")
ggsave("CorrelationB1.png", plot = correlation_plot_B1, bg = "white")

```

### Winter 2 years ago

```{r}

# combine data fames
comparison_df_2_B <- B_yearly_avg %>%
  left_join(avg_swe_2year_prior, by = "Year")

# plot
WinterSeverity_2_B <- ggplot(comparison_df_2_B, aes(x = Year)) +
  # frog abundance line
  geom_line(aes(y = Recruited_Individuals, color = "Individuals Recruited"), size = 1) +
  geom_point(aes(y = Recruited_Individuals, color = "Individuals Recruited"), size = 2) +
  # winter severity line
  geom_line(aes(y = Winter_Severity, color = "Winter Severity Two Years Prior\n(% of historical avg 1 April SWE)"), linetype = "dashed", size = 1) +
  geom_point(aes(y = Winter_Severity, color = "Winter Severity Two Years Prior\n(% of historical avg 1 April SWE)"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Individuals Recruited" = "blue", "Winter Severity Two Years Prior\n(% of historical avg 1 April SWE)" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Recruitment and Winter Severity Two Years Prior",
    x = "Year",
    y = "Value"
  )

# correlation
correlation_2_B <- cor(comparison_df_2_B$Winter_Severity, comparison_df_2_B$Recruited_Individuals)
cor_test_2_B <- cor.test(comparison_df_2_B$Winter_Severity, comparison_df_2_B$Recruited_Individuals)

correlation_plot_2_B <- ggplot(data = comparison_df_2_B, aes(x = Winter_Severity, y = Recruited_Individuals)) +
  geom_point() +
  geom_smooth(method = "lm" ,se = FALSE) +
  annotate("text", x = 75, y = 185, label = paste0("Correlation: ", round(correlation_2_B, 2)), size = 5, color = "red") +
  labs(
    title = "Correlation between Winter Severity (2-Year lag) and Recruitment",
    x = "Winter Severity (% of Historical Average 1 April SWE)",
    y = "Total Yearly Recruitment (Number of individuals)"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

print(WinterSeverity_2_B)
print(correlation_plot_2_B)

```
### Three-year rolling average no lag

```{r}

# combine data fames
comparison_df_B3 <- B_yearly_avg %>%
  left_join(avg_swe_3yrAvg, by = "Year")

# plot
WinterSeverity_B3 <- ggplot(comparison_df_B3, aes(x = Year)) +
  # frog abundance line
  geom_line(aes(y = Recruited_Individuals, color = "Individuals Recruited"), size = 1) +
  geom_point(aes(y = Recruited_Individuals, color = "Individuals Recruited"), size = 2) +
  # winter severity line
  geom_line(aes(y = swe_3yrAvg, color = "Winter Severity 3 yr avg\n(% of historical avg 1 April SWE)"), linetype = "dashed", size = 1) +
  geom_point(aes(y = swe_3yrAvg, color = "Winter Severity 3 yr avg\n(% of historical avg 1 April SWE)"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Individuals Recruited" = "blue", "Winter Severity 3 yr avg\n(% of historical avg 1 April SWE)" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Recruitment and 3-yr avg Winter Severity",
    x = "Year",
    y = "Value"
  )

# correlation
correlation_B3 <- cor(comparison_df_B3$Recruited_Individuals, comparison_df_B3$swe_3yrAvg)
cor_test_B3 <- cor.test(comparison_df_B3$Recruited_Individuals, comparison_df_B3$swe_3yrAvg)

correlation_plot_B3 <- ggplot(data = comparison_df_B3, aes(x = swe_3yrAvg, y = Recruited_Individuals)) +
  geom_point() +
  geom_smooth(method = "lm" ,se = FALSE) +
  annotate("text", x = 75, y = 185, label = paste0("Correlation: ", round(correlation_B3, 2)), size = 5, color = "red") +
  labs(
    title = "Correlation between 3-yr avg Winter Severity and Recruitment",
    x = "Winter Severity (% of Historical Average 1 April SWE)",
    y = "Yearly Mean Abundance"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

######################## Print
print(WinterSeverity_B3)
print(correlation_plot_B3)

```

### Three-year rolling average w/ 1-yr lag

```{r}

# combine data fames
comparison_df_B3.1 <- B_yearly_avg %>%
  left_join(avg_swe_3yrAvg_1yrLag, by = "Year")

# plot
WinterSeverity_B3.1 <- ggplot(comparison_df_B3.1, aes(x = Year)) +
  # frog abundance line
  geom_line(aes(y = Recruited_Individuals, color = "Individuals Recruited"), size = 1) +
  geom_point(aes(y = Recruited_Individuals, color = "Individuals Recruited"), size = 2) +
  # winter severity line
  geom_line(aes(y = swe_3yrAvg_1yrLag, color = "Winter Severity 3 yr avg 1 yr lag\n(% of historical avg 1 April SWE)"), linetype = "dashed", size = 1) +
  geom_point(aes(y = swe_3yrAvg_1yrLag, color = "Winter Severity 3 yr avg 1 yr lag\n(% of historical avg 1 April SWE)"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Individuals Recruited" = "blue", "Winter Severity 3 yr avg 1 yr lag\n(% of historical avg 1 April SWE)" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Recruitment and 3-yr avg Winter Severity with 1-yr lag",
    x = "Year",
    y = "Value"
  )

```

## Survival Probability - Winter Severity Comparison

### Same Winter

```{r}

# combine data fames
comparison_df_3 <- phi_overwinter %>%
  left_join(avg_swe, by = "Winter")

# Ensure Winter is a factor in the correct order
comparison_df_3 <- comparison_df_3 %>%
  mutate(Winter = factor(Winter, levels = c("2014-2015", "2015-2016", "2016-2017", "2017-2018"))) %>%
  mutate(Survival_Percentage = Overwinter_Survival_Probability * 100)

# Plot
WinterSeverity_Survival <- ggplot(comparison_df_3, aes(x = Winter)) +
  # survival line
  geom_line(aes(y = Survival_Percentage, color = "Survival Percentage"), size = 1, group = 1) +
  geom_point(aes(y = Survival_Percentage, color = "Survival Percentage"), size = 2) +
  # winter severity line
  geom_line(aes(y = Winter_Severity, color = "Winter Severity"), linetype = "dashed", size = 1, group = 2) +
  geom_point(aes(y = Winter_Severity, color = "Winter Severity"), size = 2) +
  # Labels and theme
  scale_color_manual(
    name = "Legend",
    values = c("Survival Percentage" = "blue", "Winter Severity" = "red")
  ) +
  theme_minimal() +
  labs(
    title = "Survival and Winter Severity",
    x = "Winter",
    y = "Value"
  )

# correlation
correlation3 <- cor(comparison_df_3$Winter_Severity, comparison_df_3$Overwinter_Survival_Probability)
cor_test3 <- cor.test(comparison_df_3$Winter_Severity, comparison_df_3$Overwinter_Survival_Probability)

correlation_plot3 <- ggplot(data = comparison_df_3, aes(x = Winter_Severity, y = Survival_Percentage)) +
  geom_point() +
  geom_smooth(method = "lm" ,se = FALSE) +
  annotate("text", x = 75, y = 100, label = paste0("Correlation: ", round(correlation3, 2)), size = 5, color = "red") +
  labs(
    title = "Correlation between Winter Severity and Survival",
    x = "Winter Severity (% of Historical Average 1 April SWE)",
    y = "Overwinter Survuval Probability"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

```