### 2025-01-31 code/R/Survival_Probability_Difference_MF.Rmd
calculates and plots M-F difference in survival

---
title: "Survival Probability Difference M-F"
output: html_document
date: "2025-01-30"
---

```{r}

library(data.table)
library(tidyverse)
library(ggplot2)

```

## Extract draws from the model with sex as a survival covariate

```{r}

# laod the model with sex as a covariate of survival
mod <- readRDS("/Users/johnimperato/Desktop/Mossy-Pond_companion/sex_survival_covariate_model.rds")

# extract the draws of the relevant variables
phi_draws <- as.data.table(mod_sex$m_fit$draws(variables = c('beta_phi', 'eps_phi'), format = 'draws_df'))

# add indexing row
phi_draws <- phi_draws %>%
  mutate(draw_id = row_number()) %>%
  select(draw_id, everything())

```

```{r}

# Identify all 'eps_phi[t]' columns using regex
eps_phi_cols <- grep("^eps_phi\\[", names(phi_draws), value = TRUE)

# Extract 'eps_phi[t]' into a separate matrix (M x T)
eps_phi_matrix <- as.matrix(phi_draws %>% select(all_of(eps_phi_cols)))

# Verify dimensions (M = number of draws, T = number of time periods)
dim(eps_phi_matrix)  # Should return c(M, T)

# Extract 'beta_phi[1]' and 'beta_phi[2]' as numeric vectors
beta_phi1_draws <- phi_draws$`beta_phi[1]`
beta_phi2_draws <- phi_draws$`beta_phi[2]`

```

```{r}

# Determine number of posterior draws (M) and time periods (T)
M <- nrow(phi_draws)
T <- ncol(eps_phi_matrix)

# Initialize matrices to store survival probabilities and their differences
female_survival <- matrix(NA, nrow = M, ncol = T)
male_survival   <- matrix(NA, nrow = M, ncol = T)
diff_survival   <- matrix(NA, nrow = M, ncol = T)

# Assign column names for clarity (optional)
colnames(female_survival) <- paste0("time_", 1:T)
colnames(male_survival)   <- paste0("time_", 1:T)
colnames(diff_survival)   <- paste0("time_", 1:T)

# Compute survival probabilities and their differences
for(m in 1:M){
  for(t in 1:T){
    # Logit-scale survival for females: beta_phi[1] + eps_phi[t]
    logit_female <- beta_phi1_draws[m] + eps_phi_matrix[m, t]

    # Logit-scale survival for males: beta_phi[1] + beta_phi[2] + eps_phi[t]
    logit_male <- beta_phi1_draws[m] + beta_phi2_draws[m] + eps_phi_matrix[m, t]

    # Convert logits to probabilities using the logistic function
    p_female <- 1 / (1 + exp(-logit_female))
    p_male   <- 1 / (1 + exp(-logit_male))

    # Store the probabilities
    female_survival[m, t] <- p_female
    male_survival[m, t]   <- p_male

    # Compute and store the difference (male - female)
    diff_survival[m, t]   <- p_male - p_female
  }
}

```

```{r}

# Define a helper function to compute summary statistics
summarize_draws <- function(x) {
  data.frame(
    mean          = mean(x),
    median        = median(x),
    sd            = sd(x),
    lower95       = quantile(x, 0.025),
    upper95       = quantile(x, 0.975),
    prob_positive = mean(x > 0)   # Proportion of draws where male > female
  )
}

# Apply the summary function to each time period's difference
time_summaries_list <- lapply(1:T, function(t) {
  summarize_draws(diff_survival[, t])
})

# Combine the summaries into a single data frame
time_summaries_df <- bind_rows(time_summaries_list, .id = "time_index") %>%
  mutate(time_period = paste0("time_", time_index)) %>%
  select(time_period, everything())

# View the summaries
print(time_summaries_df)

```

```{r}

# Plot the difference with credible intervals
ggplot(time_summaries_df, aes(x = time_period, y = mean)) +
  geom_point(color = "blue") +
  geom_errorbar(aes(ymin = lower95, ymax = upper95), width = 0.2, color = "blue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Difference in Survival Probabilities (Male - Female) Over Time",
    x = "Time Period",
    y = "Difference in Survival Probability"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}

###############################################################################
# 5. Plot the Differences with Correct Time Order
###############################################################################

# Ensure 'time_period' is an ordered factor (if not already done)
time_summaries_df <- time_summaries_df %>%
  mutate(time_period = factor(time_period, 
                              levels = paste0("time_", 1:T),  # Define the correct chronological order
                              ordered = TRUE))

# Verify the ordering
print(levels(time_summaries_df$time_period))

# Create the plot
difference_plot <- ggplot(time_summaries_df, aes(x = time_period, y = mean)) +
  geom_point(color = "blue", size = 2) +  # Increased point size for better visibility
  geom_errorbar(aes(ymin = lower95, ymax = upper95), width = 0.2, color = "blue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Difference in Survival Probabilities (Male - Female) Over Time",
    x = "Time Period",
    y = "Difference in Survival Probability"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Tilt x-axis labels for readability
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold")  # Center and style the title
  ) +
  scale_x_discrete(limits = paste0("time_", 1:T))  # Ensure correct order on x-axis

# Display the plot
print(difference_plot)

# Save the plot to a file
ggsave("difference_in_survival_probabilities.png", plot = difference_plot, width = 10, height = 6, dpi = 300)

```