### 2025-04-17 doc/Manuscript.qmd
updates to all sections

---
format: pdf
bibliography: references.bib
execute: 
  echo: true
---

# Title Title Title

John Imperato^1^

^1^School of the Environment, Yale University, New Haven, CT, 06511

## Abstract

## Introduction

*In amphibians, the what is traditionally referred to as a 'biodiversity crisis' is particualrly pronounced. Of xxx described amphibian species, xxx are declining and xxx have gone extinct. Stressors are numerous.*

*Frame the paper around biotic resilience. Amphibians are not typically considered very resilient. We have so many stories of doom and extinction adn collapse. More generally, when thinking about the future coexistense of people and other species on this planet, we tend to discount species' ability to change and evolve rapidly.*

*One example from the Sierra Nevada of California provides a cautiously hopeful example of evolved resistance to a highly-lethal pathogen that has caused numerous extinctions.* Probable mechanisms of recovery have been identified and the situation looks a lot like an evolutionary rescue scenario.

*But this example shows us that capacity might be limited for evolution to rescue vertebrates that have imited dispersal ability. Enter translocations, reintroductions, and human-assisted recovery of species we deem worthy. This is a new fronteir in amphibian conservation. A real strong example of evolutionary rescue and successful methods for people to facilitate wider scale recovery beyond what is natural.* Outcomes have previously been mixed (citations), but we have a new shining example here in Yosemite National Park that provides hope for MYLF across the rest of their range, and also a blueprint for other declining amphibian species affected by Bd.

*Recovery and resilience relies on these persistent populations, both in understanding the mechanisms and processes of persistence and recovery, and as conservation resources for human-assisted recovery strategies.*

*The MYLF example provides an exciting opportunity to study the persistent populaitons of this species. Most studies have so far been t large scale, relying on VES and mroe coarse grained observations. CMR is expensive and hard to do. Thus, CMR studies of these persistent populations are limited to sub-sections fo the species' geographic range. We took the opportunitiy to do a detailed study of one of these rare and special persistent population to contribute to scientific knowledge of amphibian resilience and for practical conservation management.*

Amphibians continue to experience unprecedented global declines, with infectious diseases emerging as primary drivers of biodiversity loss (Luedtke et al. 2023). Chytridiomycosis, caused by the fungal pathogen *Batrachochytrium dendrobatidis* (Bd), has been directly implicated in the decline or extinction of approximately 200 amphibian species across six continents (Skerratt et al. 2007). Amphibian populations exhibit highly variable responses to Bd infection; some appear relatively unaffected, others rapidly succumb and go extinct, some persist at chronically low numbers, and a few rare cases demonstrate population recovery following severe declines (Lips 2016). The mechanisms underlying persistence or recovery in Bd-affected populations remain incompletely understood and likely involve complex interactions among environmental, evolutionary, immunological, and demographic factors (Scheele et al. 2017).

The mountain yellow-legged frog species complex (*Rana sierrae* and *R. muscosa*), endemic to California's Sierra Nevada and Transverse mountain ranges, provides one of the most thoroughly documented examples of amphibian decline, persistence, and recovery in the face of Bd infection. Beginning in the 1970s, Bd began spreading across the Sierra Nevada, causing widespread extirpations of mountain yellow-legged frog populations (Vredenburg et al. 2010). By the early 2000s, these frogs had disappeared from over 90% of their historical range due to Bd impacts, compounded by habitat degradation and non-native fish introductions (Knapp 2000; Vredenburg et al. 2007). However, unlike many other amphibian species that are highly-susceptible to Bd infection, the mountain yellow-legged frog's story did not end with extinction. A small subset of populations survived their initial contact with Bd and, after decades of decline, began exhibiting increases in abundance and apparent resistance to infection (Knapp et al. 2016). These dynamics align with patterns of evolutionary rescue (Carlson et al. 2014), wherein a species rapidly acquires adaptive genetic variation that restores demographic viability in the presence of a persistent threat (Knapp et al. 2024).

While research has thoroughly described this shift from decline to recovery at broader geographic scales, fine-grained demographic studies at the individual population level remain scarce. Understanding the mechanisms and ecological processes underlying population recovery requires detailed demographic data at fine spatial and temporal resolutions. Unfortunately, such studies are time and resource-intensive, resulting in relatively few well-documented examples. Furthermore, most detailed demographic studies have concentrated in the southern portion of the species' range, primarily within or south of Yosemite National Park (Fellers et al. 2013; Joseph and Knapp 2018; Knapp et al. 2024). Relatively little is known about population dynamics in the northern Sierra Nevada (MYLF ITT 2018), an area characterized by smaller, lower-elevation frog populations occupying more heterogeneous habitats (Brown et al. 2020).

To address these knowledge gaps, we conducted a five-year capture-mark-recapture (CMR) study of a naturally recovering, Bd-persistent population of *R. sierrae* in California's Tahoe National Forest. Our objectives were to estimate demographic parameters critical for understanding population dynamics, including abundance, survival, recruitment, and detection probabilities. By quantifying these demographic parameters at a fine scale, we sought to elucidate the ecological processes supporting persistence and recovery in amphibian populations living with endemic Bd infection. Additionally, we examined how climatic variables influence these demographic parameters, as climate is expected to be a primary driver of population dynamics in these high-elevation systems where Bd has become an established component of the ecosystem (citation for climate effects).

Understanding the dynamics of these naturally recovering populations is particularly important given their significant potential to contribute to conservation efforts. Recent research has demonstrated that individuals from these populations can serve as effective founders for reintroduction programs, establishing self-sustaining populations with low long-term extinction probability when translocated to previously occupied habitats (Knapp et al. 2024). This approach leverages the apparent disease resistance that has evolved in these populations, facilitating recovery at ecologically-relevant scales. Our demographic study provides critical information to support such conservation interventions, particularly for the understudied northern Sierra Nevada populations. By quantifying vital rates and identifying factors that influence population viability in Bd-endemic environments, our results can inform selection of source populations, optimal timing of translocations, and post-translocation monitoring protocols.

## Methods

### Study area and species

The Mossy Pond study area is located in the Tahoe National Forest, north of highway 80 in Nevada County, California. It consists of approximately 60 lakes, ponds, and streams set on granite benches southeast of Fordyce Lake. Elevation ranges from 6,400 feet to 7,100 feet. The study area is a one-square-mile section of the complex, containing the namesake Mossy Pond, its seasonally-flowing outlet stream, and 12 ephemeral ponds. **\*\*Rationale for study area selection\*\*.**

The United States Geological Survey (USGS) first detected SNYLF in the complex in 1998, and the California Department of Fish and Wildlife (CDFW) has been monitoring the population since 2001.

### Quantifying Bd loads

To investigate the Mossy Pond *R. sierrae* population's Bd status, we collected epithelial swabs from 24 adult individuals over three years (2008, 2010, and 2021). Partner scientists quantified the amount of Bd DNA on each swab using real-time quantitative polymerase chain reaction (qPCR) analysis (Boyle et al. 2004). We report infection loads as the number of ITS1 (Internal Transcribed Spacer 1) gene copies detected (Longo et al. 2013).

Quantifying Bd loads rather than simply testing for Bd presence allowed us to compare our study population with other known recovering populations. Partner scientists provided *R. sierrae* Bd load data from three naturally recovering populations in Yosemite National Park for this comparison (Knapp et al. 2024).

### Frog population surveys

We studied *R. sierrae* with capture-mark-recapture (CMR) methods, which are widely used for estimating wildlife population parameters such as abundance, survival, and recruitment (Williams et al. 2001). We conducted CMR sampling according to Pollock's robust design [@pollock1982], which involves sampling at two temporal scales: primary periods (between which the population is open to demographic change) and secondary periods, during which the population is assumed closed. Each summer from 2014 to 2018, we visited the study site 3-4 times (primary periods) for three consecutive days of survey (secondary periods), for a total of 47 individual surveys.

Surveys occurred generally from 0900 to 1700, when *R. sierrae* are most active. During these hours, we continuously walked through and around all water bodies in the study area looking for frogs. We captured frogs by hand or dip net and processed them at the point of capture. We first scanned each captured frog with a BioMark 601 passive integrated transponder (PIT) tag reader (BioMark, Boise, Idaho, USA) to detect if the frog was marked (i.e., a recapture). We then used calipers to measure the snout to urostyle (SUL) length of each captured frog, and released frogs \<40 mm SUL (which are considered subadults) without further processing. **Frogs larger than 40mm (adults) required further processing.** *We continued collecting data on larger (\>40 mm SUL) frogs, which are considered adults*. For new adult captures, we inserted an 8 X 1.4 mm PIT tag (BioMark, Boise, Idaho, USA) under the dorsal skin using methods recommended by McAllister et al. (2004) [@mcallister2004]. Next, we recorded the frog's sex and weighed it inside a tared plastic bag to the nearest 0.1 g using a Pesola spring scale. Finally, we collected a GPS point (estimated precision error \~3 m) for each capture using a handheld Garmin GPS unit (Garmin citation). We only collected length and weight data during the first capture event within a three-day primary period. Afterward, if crews captured the same individual on a subsequent day during the same primary period, crews only recorded PIT tag, sex, and location data.

### Statistical Analysis

We conducted our statistical analyses using a two-stage approach that leverages Bayesian uncertainty quantification. In the first stage, we estimated survival probability, detection probability, recruitment, and adult population size for the Mossy Pond *R. sierrae* population using a site-specific open population Jolly-Seber CMR model implemented in a Bayesian framework with R statistical software \[[\@R](v4.4.2)\]. Our model was adapted from the mrmr R package [@joseph2019] (https://snarl1.github.io/mrmr/index.html), developed expressley to analyze robust design CMR data from *R. sierrae* populations. The model represents a superpopulation of M individuals whose states (not recruited, alive, dead) and observations (detected, not detected) evolve through time, with any new frogs assumed to arise from local reproduction [@knapp2024]. We assessed model convergence via trace plots and Gelman–Rubin statistics (rhat) (find a citation) and compared alternative parameterizations using leave-one-out (LOO) cross-validation (citation).

In the second stage of our statistical analysis, we leveraged the posterior distributions from our Bayesian CMR model to investigate environmental drivers of temporal variation in population parameters. For each of the resulting 4000 posterior samples, we calculated correlation coefficients between winter severity and demographic parameters (recruitment, abundance, and survival), yielding posterior distributions of these correlations. This approach allowed us to quantify both the strength of relationships between environmental factors and population dynamics and the uncertainty in these relationships, while maintaining the uncertainty structure estimated in the first stage. We ran multiple correlation analyses on these variables, lagging winter severity by 0, 1, 2, and 3 years, as well as a 3-year rolling average, to account for possible effects of the species' life history strategy, where individuals will overwinter as tadpoles 1-4 times before morphing to frogs.

*In one approach, we calculated the Pearson correlation coefficient between winter severity and recruitment for each draw, resulting in a posterior distribution of correlation coefficients that summarizes the strength and direction of the association.*

*In another approach, we performed separate linear regressions on each draw to derive a distribution of regression coefficients. Both methods allowed us to integrate the uncertainty from the capture–recapture model into our inference on the effects of environmental variables, ensuring that our conclusions about the relationship between winter severity and recruitment reflect the full uncertainty of the underlying estimates.*

### Winter severity calcualtion

We defined winter severity as the percent of historical average snow-water equivalent (SWE) remaining on 1 April of each year (100% indicating average severity). This relative metric, rather than absolute SWE depth, allowed us to incorporate data from ten snow observation stations surrounding our study area, thereby addressing the high spatial variability of snowpack distributions documented in mountainous regions (Revuelto et al., 2014). Snowpack distribution is strongly influenced by topographic position indices and variables such as elevation, aspect, and slope, which can create significant heterogeneity even within small catchments. By using multiple observation points (located between 5.1 and 19.5 km from the study area), we minimized site-specific biases that could result from topographic factors affecting individual stations.

We calculated winter severity between 2011-2018 for all ten stations, then used ArcGIS Pro (citation) to interpolate these values across the landscape using inverse-distance weighting methods, ensuring stations closer to our study area had proportionally greater influence on the final measurements. This spatial interpolation approach produced a continuous surface of winter severity values from which we extracted the average for our study area (Figure X). This methodology is appropriate for ecological studies in topographically complex regions where a single measurement station would be unlikely to represent conditions across the entire habitat of a mobile species. All snowpack data was obtained from the California Department of Water Resources via the California Data Exchange Center (CDEC—http://cdec.water.ca.gov).

## Results

### Bd status

xx / xx samples tested positive for Bd, with low to moderate infection intensities ranging from xx to xx ITS1 gene copies. These infection intensities are well below the levels associated with increased MYL frog mortality; a rough threshold of 580,000 ITS1 gene copies is considered lethal for MYL frogs, and loads reach as high as xxxxxxxxx ITS1 gene copies during epizootic events leading to extirpation (Vredenburg et al. 2010). The infection intensities observed in our study population are typical of other naturally recovering *R. sierrae* populations (figure xx), such as those now serving as translocation donors in Yosemite National Park (Knapp et al. 2024). \_\_\_\_\_\_\_ describe enzootic dynamics? (Briggs et al. 2016).

### Abundance, Survival, Recruitment, and Detection

*Abundance*

We calculated the abundance of the Mossy Pond *R. sierrae* population for each of the 16 primary survey periods of the study. The lowest abundance in the population was observed in the final period, September 2018 (117 individuals, 95% Bayesian credible interval: 103-132) and a maximum of 245 individuals (95% credible interval 237-256) was observed during the fifth primary period, in July 2015. Across all primary periods, the mean abundance of the population was 168 individuals. The population grew from 2014 to 2016 and declined in 2017 and 2018. Within years, abundance typically peaked during the second of three yearly primary periods, which was most often in July.

*Recruitment*

A total of 629 frogs recruited into the adult population over the course of the study. Annual recruitment ranged from 60 individuals (95% credible interval: 45-78) in 2018 to 166 individuals (95% credible interval: 143-188) in 2015.

*Survival*

The probability of surviving between primary periods, averaged over all frogs, was relatively high (average \_\_\_) and constant (\_\_\_), excluding one outlier period over the 2015-2016 winter (prob = 0.57).

*Detection*

For each session (secondary period; n=47) of the study, we calculated the probability that an individual frog that is alive and available for detection is actually detected. Detection probability was lowest (median 0.26, CrI 0.2-0.31) during the sixteenth session (09 September, 2015) and highest (median 60.9, CrI: 0.54-0.67) during the forty-third session (07 August, 2018). The average detection probability over 47 individual surveys was 0.454.

## Winter severity and its effects on population dynamics

Our study period coincided with high variation in winter severity at the Mossy Pond study area, with snowpack SWE ranging from 3.5% of historical average in 2015 to 150% in 2017 (Figure x). This natural variability provided an opportunity to explore if and how winter conditions influence *R. sierrae* population dynamics. Our analyses reveal a strong negative correlation between winter severity (lagged by one year) and both the number of frogs recruited into the adult population (median correlation coefficient: -0.92; credible interval: –0.70 to -0.98) and the abundance of the population (median correlation coefficient: -xx, credible interval: -xx to -xx). For example, the most severe winter in our study (2016–2017) corresponded to the lowest annual recruitment in 2018.

## Discussion

### A recovering population, back fromt he brink of extinction

Our results describe an *R. sierrae* population that is large, relatively stable, and coexisting with Bd in the environment. Its population dynamics appear to be driven primarily by yearly variation in weather rather than a devastating disease, just like populations of frog species that are not as susceptible to Bd (bunch of citations for weather driving population dynamics). This is a far cry from the raging infection intensities, mass mortality, sharp declines, and extirpations observed across the species' range over the past two decades (Vredenburg et al. 2010). **Need transition here** There is no single documented instances of an *R. sierrae* population not crashing soon after its first Bd epizootic event. Some of the last holdout Bd-naive R. sierrae populations are currently experiencing the familiar disease dynamics. This circumstantial evidence suggests overwhelmingly that our study population experienced similar disease dynamics in the not so distant past. Further, it's current size, relative stability, and low Bd infection intensities suggest that this population recovered from such a Bd-induced crash just like its counterparts in YNP. Low infection intensities in those populations are attributed to resistance to infection, likely evolved through natural selection for more resistant genotypes (Knapp et al. 2024).

The geographic novelty of our study shows that the Bd-persistence phenomenon described in this species extends beyond the southern portion of the species range into places where the species looks different. **Does the geographic distribution of resistance suggest anything about the mechanisms? Resistance is unlikely to to be predominantly environmental, as the habitat here is, by some measures, more hospitable to Bd growth (lower elevation, warmer temps...). Address compensatory recruitment? Address immune responses? Use the 2024 paper to talk about evolutionary rescue.**

### Implications for conservation management

### Climate and population viability

-   Intro to climate becoming a more important driver of population dynamics (Luedetke et al. 2023)

-   In the Sierra Nevada, the mountain range to which *R. sierrae* is endemic, past climate trends have been \_\_\_\_ and future trends are forecasted to be \_\_\_\_\_.

-   In light of our results, what does this mean?

## References

## Figures