### 2025-05-07 code/R/B.Rmd
initial commits of individual parameter extraction scripts

---
title: "B"
output: html_document
date: "2025-05-07"
---

```{r}
# ─────────────────────────────────────────────────────────────────────────────
# Recruitment (B) summaries by primary period & annual totals
# ─────────────────────────────────────────────────────────────────────────────

# 0) Load packages
# install.packages(c("dplyr","tidyr","stringr","readr"))
library(dplyr)
library(tidyr)
library(stringr)
library(readr)

# 1) Load model & survey data
mod     <- readRDS("/Users/johnimperato/Desktop/Mossy-Pond_companion/Model_2.0.rds")
surveys <- read_csv("/Users/johnimperato/Desktop/Mossy-Pond/data/survey_MossyPond.csv") %>%
             mutate(survey_date = as.Date(survey_date))

# 2) Build table of primary period dates
period_dates <- surveys %>%
  group_by(primary_period) %>%
  summarize(
    start_date = min(survey_date),
    end_date   = max(survey_date),
    .groups    = "drop"
  )

# 3) Extract all B draws (B[1] = recruitment 1→2, etc.)
b_draws_raw <- mod$m_fit$draws("B", format = "draws_df")

# 4) Pivot long & attach dates
b_draws <- b_draws_raw %>%
  pivot_longer(
    cols      = starts_with("B["),
    names_to  = "name",
    values_to = "B_raw"
  ) %>%
  mutate(
    # interval index → primary_period
    t              = as.integer(str_extract(name, "(?<=\\[)\\d+(?=\\])")),
    primary_period = t
  ) %>%
  left_join(period_dates, by = "primary_period")

# 5) Map primary_period → calendar Year
Year_map <- tibble(
  primary_period = 1:16,
  Year = c(
    rep(2014, 3),  # periods 1,2,3 → 2014
    rep(2015, 3),  # 4,5,6 → 2015
    rep(2016, 3),  # 7,8,9 → 2016
    rep(2017, 3),  # 10,11,12 → 2017
    rep(2018, 4)   # 13,14,15,16 → 2018
  )
)

b_draws <- b_draws %>%
  left_join(Year_map, by = "primary_period")

# 6) Interval‐specific summary (as before)
Recruitment_summary <- b_draws %>%
  group_by(primary_period, start_date, end_date) %>%
  summarize(
    median     = median(B_raw),
    lo025    = quantile(B_raw, 0.025),
    hi975    = quantile(B_raw, 0.975),
    cri_width = hi975 - lo025,
    .groups  = "drop"
  ) %>%
  arrange(primary_period)

# 7) Annual recruitment: sum within each posterior draw, then summarize
#    We group on the internal .draw index so covariance is preserved.
annual_B_draws <- b_draws %>%
  group_by(.draw, Year) %>%
  summarise(
    B_year = sum(B_raw),
    .groups = "drop"
  )

Yearly_Recruitment_summary <- annual_B_draws %>%
  group_by(Year) %>%
  summarise(
    median     = median(B_year),
    lo025    = quantile(B_year, 0.025),
    hi975    = quantile(B_year, 0.975),
    cri_width = hi975 - lo025,
    .groups  = "drop"
  ) %>%
  arrange(Year)

# 8) Print or export
print(Recruitment_summary)
print(Yearly_Recruitment_summary)

# Optionally save to CSV:
# write_csv(Recruitment_summary,        "Recruitment_summary.csv")
# write_csv(Yearly_Recruitment_summary, "Yearly_Recruitment_summary.csv")

```