### 2025-05-07 code/R/N.Rmd

---
title: 'N'
output: html_document
date: "2025-05-07"
---

```{r}

# ─────────────────────────────────────────────────────────────────────────────
# Abundance (N) summaries by primary period
# ─────────────────────────────────────────────────────────────────────────────

# 0) Load packages
# install.packages(c("dplyr","tidyr","stringr","readr"))
library(dplyr)
library(tidyr)
library(stringr)
library(readr)

# 1) Load model & surveys
mod     <- readRDS("/Users/johnimperato/Desktop/Mossy-Pond_companion/Model_2.0.rds")
surveys <- read_csv("/Users/johnimperato/Desktop/Mossy-Pond/data/survey_MossyPond.csv") %>%
             mutate(survey_date = as.Date(survey_date))

# 2) Build table of primary period dates
period_dates <- surveys %>%
  group_by(primary_period) %>%
  summarize(
    start_date = min(survey_date),
    end_date   = max(survey_date),
    .groups    = "drop"
  )

# 3) Extract all N draws (N[1] = period 2, N[2] = period 3, …)
n_draws_raw <- mod$m_fit$draws("N", format = "draws_df")

# 4) Pivot long & attach dates
n_draws <- n_draws_raw %>%
  pivot_longer(
    cols      = starts_with("N["),
    names_to  = "name",
    values_to = "N_raw"
  ) %>%
  mutate(
    t             = as.integer(str_extract(name, "(?<=\\[)\\d+(?=\\])")),
    primary_period = t
  ) %>%
  left_join(period_dates, by = "primary_period")

# 5) Summarize
Abundance_summary <- n_draws %>%
  group_by(primary_period, start_date, end_date) %>%
  summarize(
    median   = median(N_raw),
    lo025  = quantile(N_raw, 0.025),
    hi975  = quantile(N_raw, 0.975),
    ci_width = hi975 - lo025,
    .groups  = "drop"
  ) %>%
  arrange(primary_period)

# 6) Output
print(Abundance_summary)
# write_csv(Abundance_summary, "Abundance_summary.csv")

```