### 2025-05-07 code/R/p.Rmd

---
title: "p"
output: html_document
date: "2025-05-07"
---

```{r}

# ─────────────────────────────────────────────────────────────────────────────
# Detection (p) summaries by survey
# ─────────────────────────────────────────────────────────────────────────────

# 0) Load packages
# install.packages(c("dplyr","tidyr","stringr","readr"))
library(dplyr)
library(tidyr)
library(stringr)
library(readr)

# 1) Load model & survey data (add survey_idx for joining)
mod     <- readRDS("/Users/johnimperato/Desktop/Mossy-Pond_companion/Model_2.0.rds")
surveys <- read_csv("/Users/johnimperato/Desktop/Mossy-Pond/data/survey_MossyPond.csv") %>%
  mutate(
    survey_date = as.Date(survey_date),
    survey_idx  = row_number()
  )

# 2) Extract all p draws (p[1] … p[Jtot])
p_draws_raw <- mod$m_fit$draws("p", format = "draws_df")

# 3) Pivot to long form and attach survey metadata
p_draws <- p_draws_raw %>%
  pivot_longer(
    cols      = starts_with("p["),
    names_to  = "name",
    values_to = "p_raw"
  ) %>%
  mutate(
    survey_idx = as.integer(str_extract(name, "(?<=\\[)\\d+(?=\\])"))
  ) %>%
  left_join(surveys, by = "survey_idx")

# 4) Summarize detection probability by survey
Detection_summary <- p_draws %>%
  group_by(survey_idx, primary_period, survey_date) %>%
  summarize(
    median    = median(p_raw),
    lo025     = quantile(p_raw, 0.025),
    hi975     = quantile(p_raw, 0.975),
    ci_width  = hi975 - lo025,
    .groups   = "drop"
  ) %>%
  arrange(survey_date)

# 5) Print or export results
print(Detection_summary)
# write_csv(Detection_summary, "Detection_summary.csv")

```