### 2025-05-07 code/R/phi.Rmd

---
title: "Phi"
output: html_document
date: "2025-05-07"
---

```{r}

# ─────────────────────────────────────────────────────────────────────────────
# Interval & Seasonal Monthly Survival with Dates, CI width & Season
# ─────────────────────────────────────────────────────────────────────────────

# 0) install / load required packages
# install.packages(c("dplyr","tidyr","stringr","readr"))
library(dplyr)
library(tidyr)
library(stringr)
library(readr)

# 1) Load your model and survey data
mod <- readRDS("/Users/johnimperato/Desktop/Mossy-Pond_companion/Model_2.0.rds")

surveys <- read_csv("/Users/johnimperato/Desktop/Mossy-Pond/data/survey_MossyPond.csv") %>%
  mutate(survey_date = as.Date(survey_date))

# 2) Compute first & last survey date for each primary period
period_dates <- surveys %>%
  group_by(primary_period) %>%
  summarise(
    start_date = min(survey_date),
    end_date   = max(survey_date),
    .groups    = "drop"
  )

# 3) Build the intervals table: transition index, from→to, start/end dates, length
intervals <- period_dates %>%
  mutate(
    from       = primary_period,
    to         = lead(primary_period),
    next_start = lead(start_date)
  ) %>%
  filter(!is.na(to)) %>%
  mutate(
    delta_days   = as.numeric(difftime(next_start, end_date, units = "days")),
    delta_months = delta_days / 30.44,
    transition   = from
  ) %>%
  select(
    transition,
    from, 
    to,
    start_date,
    end_date,
    delta_days,
    delta_months
  )

# 4) Extract all raw posterior draws of overall_phi
phi_draws_raw <- mod$m_fit$draws("overall_phi", format = "draws_df")

# 5) Pivot to long form, parse the Stan index, join intervals, and standardize to monthly
phi_draws <- phi_draws_raw %>%
  pivot_longer(
    cols      = starts_with("overall_phi"),
    names_to  = "name",
    values_to = "phi_raw"
  ) %>%
  mutate(
    transition = as.integer(str_extract(name, "(?<=\\[)\\d+(?=\\])"))
  ) %>%
  left_join(intervals, by = "transition") %>%
  mutate(
    phi_monthly = phi_raw^(1 / delta_months)
  )

# 6) Interval‐specific summary: collapse down, then add CI width & season
interval_summary <- phi_draws %>%
  group_by(transition, from, to, start_date, end_date, delta_days, delta_months) %>%
  summarise(
    median   = median(phi_monthly),
    lo025  = quantile(phi_monthly, 0.025),
    hi975  = quantile(phi_monthly, 0.975),
    .groups = "drop"
  ) %>%
  # now that you have one row per transition, compute:
  mutate(
    cri_width = hi975 - lo025,
    season   = if_else(delta_months > 2, "winter", "summer")
  ) %>%
  arrange(transition)

# 7) Seasonal summaries (winter >2 mo vs. summer ≤2 mo)
season_summary <- phi_draws %>%
  mutate(
    season = if_else(delta_months > 2, "winter", "summer")
  ) %>%
  group_by(season) %>%
  summarise(
    median     = median(phi_monthly),
    lo025    = quantile(phi_monthly, 0.025),
    hi975    = quantile(phi_monthly, 0.975),
    ci_width = hi975 - lo025,
    .groups  = "drop"
  )

# 8) Print or export results
print(interval_summary)
print(season_summary)

# Optionally write tables for your paper:
# write_csv(interval_summary, "interval_survival_summary.csv")
# write_csv(season_summary,   "season_survival_summary.csv")

```